Implications of uniformly distributed, empirically informed priors for phylogeographical model selection: A reply to Hickerson et al.

### 2014-07-25 manuscripts/reply/reply-main.tex
More editing of reply ms.

\change{
Here, we discuss theoretical considerations for using empirically informed
priors for Bayesian model choice.
}
We then evaluate the empirical Bayesian model-averaging approach of
\citet{Hickerson2013} as a potential solution to the biases of \msb revealed by
\citet{Oaks2012}.
In their re-analysis of the dataset of \citet{Oaks2012}, \citet{Hickerson2013}
made an error by mixing different units of time, which makes the results
presented in their response difficult to interpret (see Supporting Information
for details).
We avoid this error, but still find their approach will often spuriously
support clustered divergence models.
Furthermore, the approach provides another means by which the method can
``escape'' models with large parameter space, which manifests in the preference
of models that exclude from consideration the true values of the model's
parameters.
\change{
Our results, show that it is difficult to choose a uniformly distributed prior
on divergence times that is broad enough to confidently contain the true values
of parameters while being narrow enough to avoid spurious support of models
with less parameter space.
}
We also explore the predictions of the potential causes of the bias proposed by
\citet{Oaks2012} and \citet{Hickerson2013}.
We show the primary cause of the method's tendency to incorrectly infer shared
divergences is the disparity in marginal likelihoods across divergence models
imposed by the uniform priors on divergence times.
We also show that sampling error is not a major contributer to the problem.
Fortunately, \citet{Oaks2014dpp} has introduced a method that uses more
flexible probability distributions to accommodate prior uncertainty in
divergence times without inhibiting the marginal likelihoods of models with
more divergence-time parameters, which increases the method's robustness and
power to detect temporal variation in divergences.

@@

In Equation \ref{eq:bayesrule}, we see that the posterior density of any state
$\myTheta{}$ of the model, is the prior density updated by the probability of
the data given $\myTheta{}$ (the likelihood of $\myTheta{}$).
The integral over the entire parameter space of the model only appears as a
normalizing constant in the denominator.
Thus, as long as the prior distribution contains the values of $\myTheta{}$
that maximize the probability of the data and the data are strongly informative
relative to the prior, the values of the parameters that maximize the posterior
distribution will be relatively robust to prior choice, even if the posterior
is technically incorrect due to using the data to inform the priors.
However, if we look at Equation~\ref{eq:bayesmodelchoice}, we see that in
Bayesian model choice it is now the \emph{marginal} likelihood of a model that
updates the prior to yield the model's posterior probability.
The integral over the entire parameter space of the likelihood weighted by the
prior probability density is no longer a normalizing constant, but rather is
what informs the posterior probability of the model from the data.
As a result, Bayesian model choice tends to be much more sensitive to priors
than parameter estimation.
Another important difference of Bayesian model choice illustrated by
Equation~\ref{eq:bayesmodelchoice} is that the value of interest, the
posterior probability of a model, is not a function of \myTheta{}, because
it is integrated out of the marginal likelihoods of the candidate models.
Thus, unlike parameter estimates, the estimated posterior probability of a
model is a single value (rather than a distribution) lacking a measure of
posterior uncertainty.

The justification for an empirical Bayesian approach to parameter estimation is
that giving the data more weight relative to the prior (i.e., using the data
twice), will often shift the peak of the estimated distribution nearer to the
true value(s) of the model's parameter(s).
However, there is often no such justification for model selection, because
unlike model parameters, the posterior probabilities of candidate models often
have no clear true values.
% There is no such justification for empirical Bayesian model choice, because
% there are no true values for the model probabilities being estimated.
Model posterior probabilities are inherently measures of our belief in the
models after our prior beliefs are updated by the data being analyzed.
This complicates the meaning of model posterior probabilities when Bayes' rule
is violated by informing priors with the same data to be analyzed.
By using the data twice, we fail to account for prior uncertainty and mislead
our posterior beliefs in the models being compared; we will be over confident
in some models and under confident in others.

>>>>

\change{
Here, we discuss theoretical considerations for using empirically informed
priors for Bayesian model choice.
}
We then evaluate the empirical Bayesian model-averaging approach of
\citet{Hickerson2013} as a potential solution to the biases of \msb revealed by
\citet{Oaks2012}.
In their re-analysis of the dataset of \citet{Oaks2012}, \citet{Hickerson2013}
made an error by mixing different units of time, which makes the results
presented in their response difficult to interpret (see Supporting Information
for details).
We avoid this error, but still find their approach will often spuriously
support clustered divergence models.
Furthermore, the approach provides another means by which the method can
``escape'' models with large parameter space, which manifests in the preference
of models that exclude from consideration the true values of the model's
parameters.
\change{
Our results, show that it is difficult to choose a uniformly distributed prior
on divergence times that is broad enough to confidently contain the true
parameter values while being narrow enough to avoid spurious support for models
with less parameter space.
}
We also explore the predictions of the potential causes of the bias proposed by
\citet{Oaks2012} and \citet{Hickerson2013}.
We show the primary cause of the method's tendency to incorrectly infer shared
divergences is the disparity in marginal likelihoods across divergence models
imposed by the uniform priors on divergence times.
We also show that sampling error is not a major contributer to the problem.
Fortunately, \citet{Oaks2014dpp} has introduced a method that uses more
flexible probability distributions to accommodate prior uncertainty in
divergence times without inhibiting the marginal likelihoods of models with
more divergence-time parameters, which increases the method's robustness and
power to detect temporal variation in divergences.

@@

In Equation \ref{eq:bayesrule}, we see that the posterior density of any state
$\myTheta{}$ of the model, is the prior density updated by the probability of
the data given $\myTheta{}$ (the likelihood of $\myTheta{}$).
The marginal likelihood of the model only appears as a normalizing constant in
the denominator.
Thus, as long as the prior distribution contains the values of $\myTheta{}$
that maximize the likelihood and the data are strongly informative
relative to the prior, the values of the parameters that maximize the posterior
distribution will be relatively robust to prior choice, even if the posterior
is technically incorrect due to using the data to inform the priors.
However, if we look at Equation~\ref{eq:bayesmodelchoice}, we see that in
Bayesian model choice it is now the \emph{marginal} likelihood of a model that
updates the prior to yield the model's posterior probability.
The integral over the entire parameter space of the likelihood weighted by the
prior density is no longer a normalizing constant, rather it is how the data
inform the posterior probability of the model.
Because the prior probability distributions placed on the model's parameters
have a strong affect on the integrated, or ``average'', likelihood of a model,
Bayesian model choice tends to be much more sensitive to priors than parameter
estimation \citep{Jeffreys1939,Lindley1957}.
Another important difference of Bayesian model choice illustrated by
Equation~\ref{eq:bayesmodelchoice} is that the value of interest, the
posterior probability of a model, is not a function of \myTheta{}, because
it is integrated out of the marginal likelihoods of the candidate models.
Thus, unlike parameter estimates, the estimated posterior probability of a
model is a single value (rather than a distribution) lacking a measure of
posterior uncertainty.

The justification for an empirical Bayesian approach to parameter estimation is
that giving the data more weight relative to the prior (i.e., using the data
twice) will often shift the peak of the estimated distribution nearer to the
true value(s) of the model's parameter(s).
However, there is often no such justification for model selection, because
unlike model parameters, the posterior probabilities of candidate models often
have no clear true values.
% There is no such justification for empirical Bayesian model choice, because
% there are no true values for the model probabilities being estimated.
Model posterior probabilities are inherently measures of our belief in the
models after our prior beliefs are updated by the data being analyzed.
This complicates the meaning of model posterior probabilities when Bayes' rule
is violated by informing priors with the same data to be analyzed.
By using the data twice, we fail to account for prior uncertainty and mislead
our posterior beliefs in the models being compared; we will be over confident
in some models and under confident in others.
